Tracing the Hercules stream around the Galaxy

### 2010-09-12 tex/response.txt

	 1. Sect 4, page 6, last paragraph: The author discusses in
	 some detail, and shows plots to illustrate, the possible
	 implications of e.g.  errors in the distance determination
	 etc. The one error that is not address explicitly in the text
	 or plots is the error on the measured radial velocities - how
	 well do we need to determine these velocities in order to be
	 able to unambiguously detect the Hercules stream? 1, 5, or 10
	 km/s? Or some other number? And what happens when V0 is
	 absolutely known or if it also has an error? Perhaps the
	 result is that the errors do no matter, but in that case this
	 should be shown and arguments to why it does not matter
	 should be given. Also, will either the RVS on board Gaia or
	 APOGEE deliver radial velocities with small enough errors to
	 detect the Hercules stream?

I added a panel to Figure 5 showing the influence of realistic
line-of-sight velocity uncertainties. The expected uncertainties for
APOGEE and HERMES are below 1 km/s; for such small uncertainties the
uncertainty-free prediction and the uncertainty-convolved prediction
are indistinguishable.

I added explicit references to the Gaia, APOGEE, and HERMES
uncertainties in the sections describing these surveys.

Uncertainty in V0 only changes the scale of the predicted velocity
distribution (since it is only used to go from vlos/V0 to the observed
vlos). This is somewhat degenerate with the dynamical R_{olr}
parameter, which is discussed at the end of the general discussion in
Section 5.

@@

	 2. A brief discussion about which tracers, e.g. dwarf stars,
	 red clump, red giant stars, to be used to test the model
	 should also be included. For example, is it necessary to
	 reach as far away from the solar neighbourhood as possible
	 (eg giants), or is the major test to get as high volume
	 density of tracers as possible (eg cool dwarfs). Such a
	 discussion would be highly useful for anyone wanting to
	 construct surveys for the very purpose of detecting
	 Hercules. This discussion does not need to be long.

This discussion is part of the new "general" discussion at the start
of section 5.

@@

I hope that by adding the general discussion and the greater detail in
discussing the specific surveys that could detect the Hercules
feature, the discussion now has the required depth.

	 Additional concerns:

	 1. On page 2, last paragraph of the introduction it is stated
	 that "propose and ... test of the bar resonance model for the
	 Hercules stream". However, this statement is not really
	 carried through in the text. It might be a good idea to
	 connect back to this statement in e.g. Sect 5 and tell the
	 reader how "testable" this model is once all the
	 uncertainties of data and shortcomings of the missions has
	 been discussed. The test and properties of the bar are also
	 mentioned in the abstract, however, there is no discussion in
	 the paper of how the properties of the bar are constrained by
	 the proposed observational test. Add this discussion, which
	 would be very interesting.

The "testibility" of the model is discussed at greater length, but I
think the whole paper makes it clear that this experiment is not a
stretch for the spectroscopic surveys or Gaia. The line-of-sight
velocity uncertainties of the spectroscopic surveys are much better
than necessary to detect the feature, and not that many stars need to
be observed to detect the feature (about 500, added to the
discussion).

The discussion at the end of Section 4 also makes it clear that none
of the nuisance or dynamical parameters change how well the model can
be tested.

I added some discussion of how the prediction could be used to
constrain the properties of the bar: The prediction depends on the
exact properties of the bar, and while some of these dependencies are
degenerate with dependecies on nuisance parameters, observing the
velocity distribution at various locations will break some of these
degeneracies. To actually calculate expected constraints would involve
simulating the data, which are very unknown for most of the proposed
surveys, but it is clear that constraints can be achieved using this
experiment.

### 2010-09-12 tex/hercules.tex
groups-simulations: finish referee edits

The velocity distribution of nearby stars contains various large
overdensities of co-moving stars that cannot be explained as the
vestiges of clumpy star formation \citep[\eg,][]{Bovy10a}. In
particular, the Hercules moving group, a group of stars significantly
offset from the bulk of the observed velocity distribution
\citep[\eg,][see \figurename~\ref{fig:obs}]{Dehnen98b,Bovy09a},
displays a wide range of ages and metallicities
\citep{Blaauw70a,raboud98a,caloi99a,Bensby07a,Bovy10a}. A compelling
model has the Hercules stream originating through resonant
interactions of stars in the outer disk with the bar in the central
region of the Galaxy \citep{dehnen00a,fux01a}, but there is no
confirmation of this picture beyond the locally observed stellar
kinematics that it was proposed to explain. An investigation of the
Hercules members' color--magnitude diagram found no evidence for any
significant metallicity anomaly with respect to other local stars
\citep{Bovy10a}, even though such a difference should be expected as
the Hercules stars originate from a few kiloparsec toward the Galactic
center where the average metallicity is higher than in the Solar
neighborhood. The absence of the predicted metallicity anomaly could
potentially be explained, \eg, through radial mixing
\citep{sellwood02a}, but it is clear that chemodynamical modeling of
the Galactic disk is not currently up to the task of separating out
these effects.

@@

The discussion in the previous sections shows that the bar-origin
paradigm for the Hercules moving group makes strong predictions about
how the Hercules feature should show up when we probe the velocity
distribution in parts of the Galaxy far removed from the Solar
neighborhood. In order to most strongly test the bar-origin for the
Hercules moving group it is best to extrapolate this picture far from
the Solar neighborhood and look at regions further than a few kpc
removed. \Gaia 's parallaxes, proper motions, and mission-averaged
line-of-sight velocities for dwarfs nearby and giants at distances
greater than a few kpc will allow the full two-dimensional velocity
distribution to be confronted with the predictions of
\figurename~\ref{fig:rphi2d} (see below). However, the final
\Gaia\ catalog will only be available around 2020.

The predicted line-of-sight velocity distributions in the previous
section show that the bar-origin theory for the Hercules moving group
can be tested with line-of-sight velocities alone. At distances
greater than a few kpc this will mean observing giants to trace the
velocity distribution. Large, planned spectroscopic surveys with
line-of-sight velocity uncertainties less than 1 km s$^{-1}$ can test
the predictions made in this paper in the next few years (see
below). These surveys will not need to observe many stars to test the
predictions laid out here: Drawing samples from the predicted velocity
distribution in \figurename~\ref{fig:1dvar} shows that about 500 stars
suffice to distinguish the bar-origin model from the axisymmetric
prediction (based on a KS test with $p < 0.01$). Since large distance
uncertainties do not affect the prediction very much (see
\figurename~\ref{fig:1dvar}), these stars can even be sampled
throughout a volume of 500 pc or more.

The bottom two panels of \figurename~\ref{fig:1dvar} show that the
predicted line-of-sight velocity distribution depends strongly on the
dynamical parameters of the bar. Changing the pattern speed of the bar
through \Rolr\ shifts the predicted distribution and in particular the
location of the Hercules feature. While this shift at a particular
location is degenerate with the shift due to the uncertainty in the
local circular velocity, results from various locations can be
combined to break this degeneracy and the local circular velocity is
also already quite well constrained \citep[\eg,][]{Bovy09b}. Changing
the strength of the bar changes the relative heights of the two main
peaks in the predicted velocity distribution. However, the main peak
of the velocity distribution might not be as simple as the one
predicted here due to other dynamical effects. Integrated measures,
such as the number of stars in the main peak and the number of stars
in the Hercules feature, could mitigate this somewhat. While
predicting actual constraints would involve detailed simulations of
the expected data, it is clear that constraints on the bar can be
derived from the experiment proposed in this paper.

@@

Since the extinction in the optical bandpasses is rather large in the
directions identified as promising in \figurename~\ref{fig:detect}
\citep[$A_V \approx 8$ mag toward $l \approx
  270^{\circ}$;][]{marshall06a}, spectroscopic surveys in the
near-infrared such as \apogee\ are the best option for seeing the
Hercules feature in the line-of-sight direction. The range of Galactic
longitudes that \apogee\ can target along the disk is $-5^{\circ} \leq
l \leq 250^{\circ}$ (SDSS-III Project Description\footnote{Available
  at \url{http://sdss3.org/collaboration/description.pdf}~.}). From
\figurename~\ref{fig:detect} this limit is right on the edge of where
we might hope to detect the Hercules feature. The Hercules bump in the
direction $l = 250^{\circ}$ is about half as strong as in
\figurename~\ref{fig:1dvar}, but remains as a clear signature of the
bar. Since \apogee\ is expected to start taking data in 2011 and to
finish operations before the first \Gaia\ data release, \apogee\ could
potentially trace the Hercules stream around the Galaxy before
\Gaia\ does. Since \apogee\ will also obtain detailed elemental
abundances for its target stars, this would also allow for a more
detailed analysis of the abundances of stars in the Hercules stream
\citep[following, \eg,][]{Bensby07a,Bovy10a}, which could shed more
light on radial mixing in the Galactic disk.

BOVY: 500 stars is enough for APOGEE to distinguish from axisymmetric
background.

\citep{Freeman10a}.

@@

@@

\clearpage
\begin{figure}
\includegraphics[width=0.35\textwidth]{distuncertain.ps}
\includegraphics[width=0.35\textwidth]{vlosuncertain.ps}\\
\includegraphics[width=0.35\textwidth]{df.ps}
\includegraphics[width=0.35\textwidth]{slope.ps}\\
\includegraphics[width=0.35\textwidth]{rolr.ps}
\includegraphics[width=0.35\textwidth]{barstrength.ps}
\caption{Variation of the predicted velocity distribution at R = 1.075
  \Ro , $\phi = -21^{\circ}$ (d= 0.4 \Ro, $l$ = 270$^{\circ}$) with
  distance uncertainties; line-of-sight velocity uncertainties;
  parameters and shape of the distribution function; shape of the
  rotation curve; bar \Rolr; and bar strength.}\label{fig:1dvar}
\end{figure}

>>>>

The velocity distribution of nearby stars contains various large
overdensities of co-moving stars that cannot be explained as the
vestiges of clumpy star formation \citep[\eg,][]{Bovy10a}. In
particular, the Hercules moving group, a group of stars significantly
offset from the bulk of the observed velocity distribution
\citep[\eg,][see \figurename~\ref{fig:obs}]{Dehnen98b,Bovy09a},
displays a wide range of ages and metallicities
\citep{Blaauw70a,raboud98a,caloi99a,Bensby07a,Bovy10a}. A compelling
model has the Hercules stream originating through resonant
interactions of stars in the outer disk with the bar in the central
region of the Galaxy \citep{dehnen00a,fux01a}, but there is no
confirmation of this picture beyond the locally observed stellar
kinematics that it was proposed to explain. An investigation of the
Hercules members' color--magnitude diagram found no evidence for any
significant metallicity anomaly with respect to other local stars
\citep{Bovy10a}, even though such a difference should be expected as
the Hercules stars originate from a few kiloparsec toward the Galactic
center where the average metallicity is higher than in the Solar
neighborhood. The absence of the predicted metallicity anomaly could
potentially be explained, \eg, through radial mixing
\citep{sellwood02a}, but it is clear that chemodynamical modeling of
the Galactic disk is not currently up to the task of separating out
these effects \citep[\eg,][]{samland03a,roskar08a}.

@@

The discussion in the previous sections shows that the bar-origin
paradigm for the Hercules moving group makes strong predictions about
how the Hercules feature should show up when we probe the velocity
distribution in parts of the Galaxy far removed from the Solar
neighborhood. In order to most strongly test the bar origin for the
Hercules moving group it is best to extrapolate this picture far from
the Solar neighborhood and look at regions further than a few kpc
removed. \Gaia 's parallaxes, proper motions, and mission-averaged
line-of-sight velocities for dwarfs nearby and giants at distances
greater than a few kpc will allow the full two-dimensional velocity
distribution to be confronted with the predictions of
\figurename~\ref{fig:rphi2d} (see below). However, the final
\Gaia\ catalog will only be available around 2020.

The predicted line-of-sight velocity distributions in the previous
section show that the bar-origin theory for the Hercules moving group
can be tested with line-of-sight velocities alone. At distances
greater than a few kpc this will mean observing giants to trace the
velocity distribution. Large, planned spectroscopic surveys with
line-of-sight velocity uncertainties less than 1 km s$^{-1}$ can test
the predictions made in this paper in the next few years (see
below). These surveys will not need to observe many stars to test the
predictions laid out here: Drawing samples from the predicted velocity
distribution in \figurename~\ref{fig:1dvar} shows that about 500 stars
suffice to distinguish the bar-origin model from the axisymmetric
prediction (based on a KS test with $P < 0.01$). Since large distance
uncertainties do not affect the prediction very much (see
\figurename~\ref{fig:1dvar}), these stars can even be sampled
throughout a volume of 500 pc or more. These spectroscopic surveys
will also obtain detailed elemental abundances for their target stars,
so they would also allow for a more detailed analysis of the
abundances of stars in the Hercules stream \citep[following,
  \eg,][]{Bensby07a,Bovy10a}, which could shed more light on radial
mixing in the Galactic disk.

The last two panels of \figurename~\ref{fig:1dvar} show that the
predicted line-of-sight velocity distribution depends strongly on the
dynamical parameters of the bar. Changing the pattern speed of the bar
through \Rolr\ shifts the predicted distribution and in particular the
location of the Hercules feature. While this shift at a particular
location is degenerate with the shift due to the uncertainty in the
local circular velocity, results from various locations can be
combined to break this degeneracy and the local circular velocity is
also already quite well constrained \citep[\eg,][]{Bovy09b}. Changing
the strength of the bar changes the relative heights of the two main
peaks in the predicted velocity distribution. However, the main peak
of the velocity distribution might not be as simple as the one
predicted here due to other dynamical effects. Integrated measures,
such as the number of stars in the main peak and the number of stars
in the Hercules feature, could mitigate this somewhat. While
predicting actual constraints would involve detailed simulations of
the expected data, it is clear that constraints on the dynamical
properties of the bar can be derived from the experiment proposed in
this paper.

@@

Since the extinction in the optical bandpasses is rather large in the
directions identified as promising in \figurename~\ref{fig:detect}
\citep[$A_V \approx 8$ mag toward $l \approx
  270^{\circ}$;][]{marshall06a}, spectroscopic surveys in the
near-infrared such as \apogee\ are the best option for seeing the
Hercules feature in the line-of-sight direction. The range of Galactic
longitudes that \apogee\ can target along the disk is $-5^{\circ} \leq
l \leq 250^{\circ}$ and its expected line-of-sight uncertainties are
about 0.5 km s$^{-1}$ (SDSS-III Project Description\footnote{Available
  at \url{http://sdss3.org/collaboration/description.pdf}~.}). From
\figurename~\ref{fig:detect} this longitude limit is right on the edge
of where we might hope to detect the Hercules feature. The Hercules
bump in the direction $l = 250^{\circ}$ is about half as strong as in
\figurename~\ref{fig:1dvar}, but remains as a clear signature of the
bar. As before, about 500 stars a few kpc away are enough to detect
the Hercules feature. Since \apogee\ is expected to start taking data
in 2011 and to finish operations before the first \Gaia\ data release,
\apogee\ could potentially trace the Hercules stream around the Galaxy
before \Gaia\ does.

Another planned spectroscopic survey is
\hermes\ \citep{Freeman10a}. While operating in the optical, it
expects to survey hundreds of thousands of giants in the inner Galaxy,
many as far as 6 kpc away, with line-of-sight velocity uncertainties
less than 1 km s$^{-1}$ (K. Freeman, private communication). Since
\hermes\ operates from the Southern hemisphere it can observe the most
promising directions in \figurename~\ref{fig:detect} and, like
\apogee, detect the Hercules feature before \Gaia.

@@

\bibitem[Ro{\v s}kar \etal(2008)]{roskar08a}
  Ro{\v s}kar,~R., \etal, 2008,
  \apjl, 684, L79

\bibitem[Samland \& Gerhard(2003)]{samland03a}
  Samland,~M.~\& Gerhard,~O.~E., 2003,
  \aap, 399, 961

@@

\clearpage
\begin{figure}
\includegraphics[width=0.32\textwidth]{distuncertain.ps}
\includegraphics[width=0.32\textwidth]{vlosuncertain.ps}
\includegraphics[width=0.32\textwidth]{df.ps}\\
\includegraphics[width=0.32\textwidth]{slope.ps}
\includegraphics[width=0.32\textwidth]{rolr.ps}
\includegraphics[width=0.32\textwidth]{barstrength.ps}
\caption{Variation of the predicted velocity distribution at R = 1.075
  \Ro , $\phi = -21^{\circ}$ (d= 0.4 \Ro, $l$ = 270$^{\circ}$) with
  distance uncertainties; line-of-sight velocity uncertainties;
  parameters and shape of the distribution function; shape of the
  rotation curve; bar \Rolr; and bar strength.}\label{fig:1dvar}
\end{figure}
